We present the proposed decentralized fault-tolerant controller, designed to solve the property certification problem in \cref{def:problem} for the reduced-order model in \cref{sec:reduction-and-reference}. Each drone $i$ runs a single instance of the \textit{DecentralizedLocalController}, realized as a continuous-time Drake \texttt{LeafSystem<double>}. The controller is a three-layer cascade consisting of an outer-loop proportional-derivative law with a passive anti-swing slot shift, a single-step convex quadratic program that enforces the tilt-and-thrust actuator envelope while absorbing the feed-forward tension, and an inner-loop Euler-angle attitude PD; all three layers operate in continuous time. The central design element is the identity $T_i^{\mathrm{ff}} = T_i$: setting the thrust feed-forward equal to the drone's own measured rope tension encodes passive fault tolerance without fault detection, peer communication, or supervisory reconfiguration. The no-detection, no-communication property is a property of \emph{this baseline controller}; the optional reshape supervisor of \cref{sec:reshape-supervisor} adds a local tension-latch fault detector and a per-fault broadcast channel and therefore lies outside the baseline information pattern, as we make explicit there. Among the three baseline layers, only the optional $L_1$-adaptive altitude augmentation is detailed in \cref{sec:extensions}, which introduces a discrete sub-system, updated at a fixed period $\Delta t_{L_1} \equiv T_S = \SI{0.2}{\milli\second}$. The alias $T_s$ is used throughout the stability analysis of the adaptive update in \cref{sec:l1-augmentation}; its injection point in the cascade is identified within \cref{sec:outer-loop}.

\input{Figures/diagrams/fig_controller_cascade.tex}

\begin{table*}[htbp]
  \centering
  \caption{Component-level information pattern. The baseline cascade and the $L_1$ and MPC extensions consume no fault flag and no peer message; only the optional reshape supervisor steps outside this pattern.}
  \label{tab:component-pattern}
  \small
  \begin{tabular}{p{4.5cm}cccp{3.0cm}}
    \toprule
    Component & Detection? & Communication? & In main theorem? & Purpose \\
    \midrule
    Baseline cascade + identity \eqref{eq:Tff} & No & No & Yes & Passive fault absorption \\
    $L_1$ altitude augmentation & No & No & Preserves theorem & Parameter mismatch \\
    MPC tension ceiling & No & No & Preserves theorem & Tension constraint \\
    Reshape supervisor & Yes (latch) & Yes ($\lceil\log_2 N\rceil$ bits) & Optional only & Slot reassignment \\
    \bottomrule
  \end{tabular}
\end{table*}

\subsection{Controller Input--Output Signature}
\label{sec:ctrl-signature}

The controller input--output signature is as follows. At each continuous-time evaluation, drone $i$'s controller consumes the drone's own inertial state $(\vect p_i, \vect v_i, \mat R_i, \vect\omega_i^{\mathcal{W}})$ extracted from the plant state vector, the drone's own scalar rope-tension measurement $T_i(t) \ge 0$, the locally observed payload velocity $\vect v_L(t)$, the pre-distributed reference waypoints $(\vect p_L^d(t), \vect v_L^d(t))$ linearly interpolated between samples as described in \cref{sec:reduction-and-reference}, and optionally a slot-offset override $\boldsymbol\delta_i^{\mathrm{ovr}}(t)$ from a \textit{FormationCoordinator} port. When the coordinator port is disconnected, the static formation offset $\boldsymbol\delta_i$ from the parameter struct is used in place of the override. The payload position $\vect p_L$ is not accessed by the baseline controller. The information consumed is a strict subset of the admissible set $\mathcal{I}_i(t)$ defined in \eqref{eq:info-set}; no peer state, no peer-cable tension, and no fault signal are accessed. The controller emits a world-frame spatial wrench $(\vect F_i^{\mathcal{W}}, \vect\tau_i^{\mathcal{W}})$ applied to the drone body by the plant.

% ----------------------------------------------------------------
\subsection{Outer Loop: PD with Anti-Swing Slot Shift}
\label{sec:outer-loop}
% ----------------------------------------------------------------

\paragraph{Slot reference and anti-swing shift.} Each drone is pre-assigned a hovering slot displaced vertically above the payload reference point and laterally by a fixed formation offset,
\begin{equation}
  \vect p_{\mathrm{slot},i}(t)
    \;=\; \vect p_L^d(t) + \boldsymbol\delta_i
        + d_{\mathrm{rope}}\,\hat{\vect e}_3,
  \label{eq:slot-pos}
\end{equation}
where $\boldsymbol\delta_i \in \R^3$ is the formation offset (overridable at run time by the coordinator port), $d_{\mathrm{rope}} = \SI{1.25}{\metre}$ is the vertical slot elevation matching the nominal rope length $L$, and $\hat{\vect e}_3$ is the world-frame upward unit vector. The slot reference velocity is the payload reference velocity, $\vect v_{\mathrm{slot},i}(t) = \vect v_L^d(t)$. To damp the payload pendulum passively, we shift the slot position relative to the payload's horizontal velocity. Let $\vect v_{L,\perp} \triangleq [v_{L,x}, v_{L,y}, 0]^{\top}$; the horizontal swing correction is
\begin{equation}
  \vect s_i(t) =
    \begin{cases}
      -k_{\mathrm{swing}}\,\vect v_{L,\perp}
        & \text{if }
          k_{\mathrm{swing}}\|\vect v_{L,\perp}\| \le s_{\max}, \\[2pt]
      -s_{\max}\,\hat{\vect v}_{L,\perp}
        & \text{otherwise,}
    \end{cases}
  \label{eq:swing-shift}
\end{equation}
where $\hat{\vect v}_{L,\perp}$ is the unit vector of $\vect v_{L,\perp}$. The dynamically adjusted slot position is $\vect p_{\mathrm{slot},i}^{\mathrm{dyn}}(t)  = \vect p_{\mathrm{slot},i}(t) + \vect s_i(t)$.

\paragraph{Tracking errors.} The position and velocity errors are
\begin{equation}
  \vect e_p(t) = \vect p_{\mathrm{slot},i}^{\mathrm{dyn}}(t)
                 - \vect p_i(t),
  \qquad
  \vect e_v(t) = \vect v_{\mathrm{slot},i}(t) - \vect v_i(t).
  \label{eq:slot-error}
\end{equation}
The velocity error is referenced against the unshifted slot velocity; the anti-swing correction modifies the position reference only, and its time derivative does not enter the velocity error term.

\paragraph{Target acceleration.} The outer loop computes a target body acceleration,
\begin{align}
\label{eq:a-target}
  \vect a_{\mathrm{target}}(t)
    =& \underbrace{%
        \begin{bmatrix}
          K_p^{xy}\, e_{p,x} + K_d^{xy}\, e_{v,x} \\[2pt]
          K_p^{xy}\, e_{p,y} + K_d^{xy}\, e_{v,y} \\[2pt]
          K_p^{z}\;\; e_{p,z} + K_d^{z}\;\; e_{v,z}
        \end{bmatrix}}_{\vect a_{\mathrm{track}}}\\
    &+\; w_{\mathrm{swing}}
      \begin{bmatrix}
        -k_{\mathrm{swing}}\, v_{L,x} \\[2pt]
        -k_{\mathrm{swing}}\, v_{L,y} \\[2pt]
        0
      \end{bmatrix}.\nonumber
\end{align}
The altitude gain $K_p^z = 100$ substantially exceeds the horizontal gain $K_p^{xy} = 30$. The magnitude is selected so that the closed-loop altitude matrix $A_m = \bigl[\begin{smallmatrix} 0 & 1 \\ -K_p^z & -K_d^z \end{smallmatrix}\bigr]$ is Hurwitz with a dominant pole at $s_1 = -5.37\,\mathrm{s}^{-1}$ and the associated Lyapunov matrix $P_v$ (Lemma~\ref{lem:altitude-iss}) yields the decay rate $\lambda$ of \cref{thm:hybrid-stability}. The horizontal gain $K_p^{xy} = 30$ stabilizes the horizontal channels independently and contributes to the slower pole $\alpha_{xy} = 2.37\,\mathrm{s}^{-1}$ that governs the between-fault Lyapunov contraction rate in Lemma~\ref{lem:dwell-decay}. The rope-compliance intuition that vertical rope forces dominate is the physical reason the altitude channel is assigned a higher gain; the Lyapunov-matrix derivation of Lemma~\ref{lem:altitude-iss} is the formal reason the specific value $K_p^z = 100$ is chosen rather than any other comparable value. When the $L_1$ adaptive extension is active, the adaptive altitude correction $u_{\mathrm{ad}}(t)$, computed by the discrete $\Delta t_{L_1}$-periodic sub-system, is added to the $z$-component of $\vect a_{\mathrm{target}}$ at this point in the cascade --- after the outer PD and before the QP --- leaving the horizontal channels unmodified (see \cref{sec:extensions}).

% ----------------------------------------------------------------
\subsection{Actuator Projection and Feed-Forward Tension}
\label{sec:qp-and-ff}
% ----------------------------------------------------------------

\paragraph{Single-step convex QP.} The target $\vect a_{\mathrm{target}}$ is not guaranteed to lie inside the tilt-and-thrust actuator envelope. We project it onto the feasible set through a single-step convex QP that also regularises against unnecessarily large commands:
\begin{subequations}
\label{eq:qp}
\begin{align}
  \vect a_d^\star
    &= \operatorname*{argmin}_{\vect a_d \in \R^3}
      \; w_t\,\bigl\|\vect a_d - \vect a_{\mathrm{target}}\bigr\|^2
      + w_e\,\|\vect a_d\|^2,
    \label{eq:qp-cost} \\
  \text{s.t.}\;\;
    &\abs{a_{d,x}},\;\abs{a_{d,y}}
      \le a_{\mathrm{tilt}} \triangleq g\tan\theta_{\max},
    \label{eq:qp-tilt} \\
    &a_{z,\min}(t) \le a_{d,z} \le a_{z,\max}(t),
    \label{eq:qp-vert}
\end{align}
\end{subequations}
where the vertical bounds incorporate the rope tension
feed-forward,
\begin{align}
\label{eq:az-bounds}
  a_{z,\min}(t) &= \frac{f_{\min} - T_i^{\mathrm{ff}}(t)}{m_i} - g,\\
  a_{z,\max}(t) &= \frac{f_{\max} - T_i^{\mathrm{ff}}(t)}{m_i} - g.\nonumber
\end{align}
At canonical parameters,
$a_{\mathrm{tilt}} = 9.81\tan(0.6)
\approx \SI{6.71}{\metre\per\second\squared}$.

\begin{remark}[Regularized projection]
The effort weight $w_e = 0.02$ makes \eqref{eq:qp} a regularized projection rather than a pure nearest-point projection. When all constraints \eqref{eq:qp-tilt}--\eqref{eq:qp-vert} are inactive, the unconstrained minimiser of \eqref{eq:qp-cost} is
\begin{equation}
\label{eq:qp-unconstrained}
  \vect a_d^\star
    = \frac{w_t}{w_t + w_e}\,\vect a_{\mathrm{target}}
    \approx 0.98\,\vect a_{\mathrm{target}},
\end{equation}
producing a ${\approx}2\,\%$ shrinkage relative to the target. The effort term guards against large actuator commands in low-error regimes; its effect on closed-loop performance is negligible under nominal operation.
\end{remark}

\paragraph{Feasibility and active-set regime.} The vertical interval in \eqref{eq:qp-vert} is non-empty whenever $f_{\min} < f_{\max}$, since both bounds shift by the same feed-forward term $T_i^{\mathrm{ff}}$ and the interval width $(f_{\max} - f_{\min})/m_i$ is preserved. Hover feasibility additionally requires the commanded thrust $f_i$ --- inclusive of $T_i^{\mathrm{ff}}$ --- to remain in $[f_{\min}, f_{\max}]$; in nominal operation $T_i \ll f_{\max} = \SI{150}{\newton}$ per drone, so hover feasibility holds throughout all simulated scenarios. The stability analysis of \cref{sec:stability} admits the QP under Doctrine D1: within each inter-fault window, the active set of constraints is locally constant, and transitions are absorbed into the slack-excursion budget of $\Omega_\tau^{\mathrm{dwell}}$ (see \cref{sec:Problem_Statement} for the formal statement and Table~\ref{tab:domain-audit} for empirical support at $0.10\%$ worst transition fraction). The baseline controller itself switches regimes freely as the QP's active set evolves; Doctrine D1 is a property of the analysis domain, not of the controller. If the QP solver returns infeasible, the controller falls back to a component-wise clamping of the proportional-derivative tracking component $\vect a_{\mathrm{track}}$ (the first term of \eqref{eq:a-target}, without the swing term or effort regularization) to the bounds \eqref{eq:qp-tilt}--\eqref{eq:qp-vert}. Solver failure is flagged internally for logging. We note that infeasibility has not been observed in any simulation campaign.

\paragraph{Feed-forward tension and three-phase logic.} The feed-forward tension $T_i^{\mathrm{ff}}$ follows a three-phase logic keyed to the cable-engagement state. In Phase 0 (pre-engagement), before the drone's cable engages the payload, the measured tension satisfies $T_i \approx 0$ and $T_i^{\mathrm{ff}} = T_i \approx 0$, and no thrust correction is applied. In Phase 1 (pickup ramp), cable engagement is detected when the measured tension first exceeds the threshold $T_{\mathrm{thresh}} = \SI{0.3}{\newton}$; this event latches the pickup start time $t_{\mathrm{pickup}}$. During the subsequent window $[t_{\mathrm{pickup}}, t_{\mathrm{pickup}} + T_{\mathrm{ramp}}]$, the feed-forward is bounded above by a $C^1$ smoothstep ceiling,
\begin{align}
\label{eq:Tff-ramp}
  T_i^{\mathrm{ff}}(t)
    &= \min\!\bigl(T_i(t),\;\alpha(\tau)\,\bar{T}\bigr),\\
  \tau &= \operatorname{clamp}\!\!\left(
    \frac{t - t_{\mathrm{pickup}}}{T_{\mathrm{ramp}}},\, 0,\, 1
  \right),\nonumber
\end{align}
where $\alpha(\tau) = 3\tau^2 - 2\tau^3$ is the Hermite cubic smoothstep, $\bar{T} = m_L g / N$ is the nominal per-drone tension share recomputed at simulator construction, and $T_{\mathrm{ramp}} = \SI{1.0}{\second}$. For the canonical campaigns ($m_L = \SI{10}{\kilo\gram}$, $N = 5$), $\bar{T} \approx \SI{19.62}{\newton}$. In Phase 2 (post-ramp / canonical), once the ramp completes, or immediately when the simulator option \texttt{initial\_pretensioned} is \texttt{true} (the default for all canonical simulation campaigns, which bypasses Phase 0 and Phase 1 entirely), the feed-forward reduces to the identity
\begin{equation}
  \boxed{T_i^{\mathrm{ff}}(t) \;=\; T_i(t).}
  \label{eq:Tff}
\end{equation}
The simulator flag \texttt{disable\_tension\_ff} (default \texttt{false}) overrides \eqref{eq:Tff} to $T_i^{\mathrm{ff}} = 0$; this ablation hook isolates the causal contribution of the identity \eqref{eq:Tff} in the feed-forward ablation campaign.

\begin{remark}[Passive fault tolerance via \eqref{eq:Tff}]
\label{rem:passive-ft}
The identity \eqref{eq:Tff} is necessary, not merely convenient, for the dwell-cycle Lyapunov contraction $\rho \in (0, 1)$ proved in Lemma~\ref{lem:dwell-decay} of \cref{sec:stability}. The bounded fault-jump condition of Lemma~\ref{lem:jump-continuity} keeps the post-fault overshoot of leading order in the redistribution magnitude $\Delta_f$: drone $i$'s thrust tracks the new measured tension within the rope timescale, so the altitude-channel contribution to the overshoot stays of order $\Delta_f$ rather than inflating by the uncompensated matched disturbance. Without \eqref{eq:Tff}, the altitude channel sees the tension redistribution as an unrejected matched disturbance, and the overshoot grows to a level that the dwell-cycle decay cannot absorb within one pendulum period. The feed-forward ablation campaign confirms this empirically: disabling \eqref{eq:Tff} increases post-fault RMS tracking error by $34$--$39\%$ and inflates the payload altitude-sag envelope by $3.6$--$4.0\times$ across V3, V4, and V5.
\end{remark}

\subsection{Inner Loop: Attitude PD and Controller Parameters}
\label{sec:inner-loop}

\paragraph{Thrust and desired attitude.}
Given the QP solution $a_{d,z}^\star$, the commanded thrust is
\begin{align}
\label{eq:thrust}
  f_i = \operatorname{clamp}\!\Bigl(
      m_i\bigl(g + a_{d,z}^\star\bigr)
      &+ T_i^{\mathrm{ff}}\\
      &+ \mathbf{1}_{\mathrm{ramp}}\,
        K_{p,T}\!\bigl(\alpha(\tau)\,\bar{T} - T_i\bigr),\;\nonumber\\
      &f_{\min},\; f_{\max}
    \Bigr),\nonumber
\end{align}
where $\mathbf{1}_{\mathrm{ramp}} = 1$ during Phase 1 only. The pickup-phase correction $K_{p,T}(\alpha(\tau)\bar{T} - T_i)$, with gain $K_{p,T} = 1.5$, drives the measured tension toward the ramp ceiling; it is inactive in all canonical simulations because \texttt{initial\_pretensioned = true} bypasses Phase 1. Under the small-angle approximation, the horizontal acceleration commands map to desired pitch and roll angles,
\begin{align}
\label{eq:attitude-des}
  \theta_{\mathrm{pitch}}^d
    &= \operatorname{sat}\!\!\left(\frac{a_{d,x}^\star}{g},\;\pm\theta_{\max}\right),\\
  \theta_{\mathrm{roll}}^d
    &= \operatorname{sat}\!\!\left(-\frac{a_{d,y}^\star}{g},\;\pm\theta_{\max}\right).\nonumber
\end{align}
In the ZYX Euler convention with the Z-up world frame, a positive $x$-acceleration demand is realized by a positive pitch angle $\theta_{\mathrm{pitch}}^d > 0$, which tilts the thrust vector toward $+x_{\mathcal{W}}$; a positive $y$-acceleration demand requires a negative roll angle (right wing up, i.e.\ $\theta_{\mathrm{roll}}^d < 0$), which tilts the thrust vector toward $+y_{\mathcal{W}}$.

\paragraph{Euler-angle attitude PD and output wrench.} Roll and pitch are extracted from $\mat R_i$ via the ZYX Euler decomposition, using 1-indexed matrix entries:
\begin{equation}
  \phi_{\mathrm{roll}} = \operatorname{atan2}(R_{32},\, R_{33}),
  \qquad
  \theta_{\mathrm{pitch}} = \operatorname{asin}(-R_{31}).
  \label{eq:euler-extract}
\end{equation}
Yaw is not extracted as a scalar Euler angle. Instead, the $(1,2)$-skew component of $\mat R_i$ provides a yaw error consistent with the ZYX representation,
\begin{equation}
  \epsilon_\psi \;\triangleq\; \tfrac{1}{2}(R_{21} - R_{12}),
  \label{eq:yaw-skew}
\end{equation}
which equals zero when $\mat R_i$ has zero yaw under the ZYX convention. The body-frame angular velocity is recovered by $\vect\omega_i^{\mathcal{B}} = \mat R_i^{\top}\, \vect\omega_i^{\mathcal{W}}$. The body-torque commands are
\begin{align}
  \tau_{i,x}
    &= K_p^{\mathrm{att}}\!\left(
         \theta_{\mathrm{roll}}^d - \phi_{\mathrm{roll}}
       \right)
      - K_d^{\mathrm{att}}\,\omega_{i,x}^{\mathcal{B}},
    \label{eq:att-pd-roll} \\[2pt]
  \tau_{i,y}
    &= K_p^{\mathrm{att}}\!\left(
         \theta_{\mathrm{pitch}}^d - \theta_{\mathrm{pitch}}
       \right)
      - K_d^{\mathrm{att}}\,\omega_{i,y}^{\mathcal{B}},
    \label{eq:att-pd-pitch} \\[2pt]
  \tau_{i,z}
    &= -\frac{K_p^{\mathrm{att}}}{2}\,(R_{21} - R_{12})
          - K_d^{\mathrm{att}}\,\omega_{i,z}^{\mathcal{B}},
    \label{eq:att-pd-yaw}
\end{align}
each saturated to $[-\tau_{\max}, \tau_{\max}]$. The law \eqref{eq:att-pd-roll}--\eqref{eq:att-pd-yaw} uses Euler-angle error directly rather than a geodesic $\SOthree$ error map; it is valid under the small-tilt operating regime enforced by the QP constraint \eqref{eq:qp-tilt}. The controller emits the world-frame wrench
\begin{equation}
  \vect F_i^{\mathcal{W}} = f_i\,\mat R_i\,\hat{\vect e}_3,
  \qquad
  \vect\tau_i^{\mathcal{W}} = \mat R_i\,
    [\tau_{i,x},\; \tau_{i,y},\; \tau_{i,z}]^{\top},
  \label{eq:wrench}
\end{equation}
applied to the drone body by the plant at the same continuous-time evaluation instant as all upstream computations.

\paragraph{Canonical controller parameters.} The canonical controller parameters used throughout all simulation campaigns are collected in Table~\ref{tab:baseline-params}.

\begin{table}[t]
  \centering
  \caption{Proposed controller parameters (canonical values used
  in all simulation campaigns).}
  \label{tab:baseline-params}
  \small
  \renewcommand{\arraystretch}{1.05}
  \begin{tabular}{lrl}
    \toprule
    Parameter & Value & Role \\
    \midrule
    $K_p^{xy},\; K_d^{xy}$
      & $30,\;15$
      & horizontal outer loop \\
    $K_p^{z},\;\; K_d^{z}$
      & $100,\;24$
      & altitude outer loop \\
    $K_p^{\mathrm{att}},\; K_d^{\mathrm{att}}$
      & $25,\;4$
      & attitude inner loop \\
    $k_{\mathrm{swing}}$
      & $0.8$
      & anti-swing gain \\
    $w_{\mathrm{swing}}$
      & $0.3$
      & anti-swing weight in $\vect a_{\mathrm{target}}$ \\
    $s_{\max}$
      & \SI{0.3}{\metre}
      & slot-shift saturation \\
    $d_{\mathrm{rope}}$
      & \SI{1.25}{\metre}
      & slot $z$-offset (nominal rope length) \\
    $w_t,\; w_e$
      & $1.0,\;0.02$
      & QP tracking / effort weights \\
    $\theta_{\max}$
      & \SI{0.6}{\radian}
      & tilt ceiling (${\approx}34^{\circ}$) \\
    $f_{\min},\; f_{\max}$
      & $0,\;\SI{150}{\newton}$
      & thrust envelope \\
    $\tau_{\max}$
      & \SI{10}{\newton\metre}
      & torque saturation \\
    $T_{\mathrm{thresh}}$
      & \SI{0.3}{\newton}
      & pickup-detection threshold \\
    $\bar{T}$
      & \SI{19.62}{\newton}
      & nominal per-drone tension ($m_L g / N$) \\
    $K_{p,T}$
      & $1.5$
      & pickup-phase tension correction gain \\
    $T_{\mathrm{ramp}}$
      & \SI{1.0}{\second}
      & pickup smoothstep window \\
    $\Delta t_{L_1}$
      & \SI{0.2}{\milli\second}
      & $L_1$ discrete update period \\
    \bottomrule
  \end{tabular}
\end{table}

The three-layer controller specified above is the baseline on which \cref{sec:stability} establishes the hybrid practical-ISS theorem. The only structurally nonstandard element is the feed-forward identity $T_i^{\mathrm{ff}} = T_i$ in \eqref{eq:az-bounds} and \eqref{eq:thrust}; every other block is conventional. The next section shows that this single identity is the mechanism responsible for the per-fault-cycle Lyapunov contraction, with $\rho \in (0,1)$, and that without it, the closed loop loses its hybrid practical-ISS property.
