In this paper, we addressed cooperative aerial manipulation under unannounced cable severance through a single architectural commitment: the post-fault tension redistribution reaches each surviving drone through its own actuator-side measurement on the plant's native timescale. The per-drone altitude feed-forward identity $T_i^{\mathrm{ff}}=T_i$ converts that signal directly into thrust at the next control tick. We presented a three-layer decentralized cascade --- outer-loop PD with anti-swing slot shift, single-step convex QP on the tilt-and-thrust envelope, and geometric attitude inner loop --- in which this identity is the sole non-standard element. This commitment is supported by three constructive certificates: a taut-cable reduction theorem that bounds the closed-loop deviation between the distributed-rope and lumped-rope plants by $O(\delta+\eta_{\max})$ on a slack-excursion-bounded admissibility set; a hybrid practical-ISS recovery envelope that yields a closed-form per-cycle Lyapunov contraction $\rho\in(0,1)$ across each unannounced severance; and a closed-form discrete-time stability bound that certifies the optional $L_1$ adaptive altitude augmentation, removing the dominant hyper-parameter of the architecture from empirical tuning.

We validated the proposed controller in high-fidelity Drake multibody simulation through a six-variant capability sequence and four targeted ablation campaigns at $N=5$, $m_L=10\,\mathrm{kg}$, and $4\,\mathrm{m\,s^{-1}}$ Dryden wind. The feed-forward ablation pins the identity as the causal recovery mechanism: removing it inflates cruise RMSE by $34$--$39\,\%$ and post-fault sag by $3.6$--$4.0\times$, while the intact controller absorbs each severance through a thrust spike that settles within one payload-pendulum period and never reaches the saturation hazard line. The measured per-cycle Lyapunov contraction sits well below the $\rho<1$ sufficiency bound and remains below unity even at sub-threshold dwell. The per-drone information set carries no peer state and no fault flag; the recovery envelope therefore extends by induction to $F\le N{-}2$ as long as actuator margin remains, and the demonstrated configuration consumes only a fraction of that margin. The architecture eliminates the announcement-detection-reconfiguration timeline entirely: in this regime, the identity \emph{is} the response.

For future work, we identify four directions within the architectural category. First, a parallel reduced-rope simulator that audits the reduction theorem at the trajectory level rather than at the domain-gate level. Second, a multi-fault generalization of the dwell-time argument beyond pairwise composition, informed by the empirical sub-threshold contraction we observed. Third, a tube-MPC tightening that folds the post-fault overshoot of the recovery envelope into the constraint right-hand side, converting the spike-dominated null finding into a positive result on trajectories where the dynamic spike is the binding constraint. Fourth, a head-to-head comparison against active fault-tolerant baselines at the actuator-tick rate, isolating the detection-loop contribution to recovery latency. Drone-side faults, soft cable failures, tension-sensor failures, and payload-attitude dynamics lie outside the architectural category; each requires its own announcement channel and is properly the subject of a different paper. The cable-severance instance is the certification vehicle, not the contribution. The contribution is the category itself: for self-announcing structural transitions on the plant's native timescale, the announcement is the response, and detection is unnecessary.