%We are interested in cooperative aerial transport, in which $N$ multirotor uncrewed aerial vehicles (UAVs) lift a common payload via flexible cables~\cite{Michael2011, Fink2011, Bernard2011, Tagliabue2019}. Cable severance---from abrasion, hardware fatigue, or obstacle contact---is a structural fault: it discretely reduces the constraint set, redistributes load across surviving cables within one payload-pendulum period ($\tau_{\mathrm{pend}} = \SI{2.24}{\second}$ at $L = \SI{1.25}{\metre}$), and requires a compensatory response within that window.

%Unilateral coupling, hybrid constraint-cardinality transitions at severance events~\cite{GoebelSanfeliceTeel2012, Brogliato2016}, and strict information locality together rule out classical cooperative and slung-load architectures: any admissible controller must deliver post-fault response without detection, annunciation, or reconfiguration and must remain analytically valid across structural transitions.

%This paper shows that, for this plant class, a single feed-forward identity in the per-drone controller can provide the dominant passive fault-response mechanism, rather than relying on detection and reconfiguration: drone $i$'s altitude thrust feed-forward is set equal to its own measured rope tension,
%\begin{equation}
%  T_i^{\mathrm{ff}}(t) = T_i(t).
%\end{equation}
%While the use of measured interaction forces in control is not new, we show that, in this setting, this identity has a non-obvious consequence: it induces a certifiable stabilizing response across structural fault transitions on the plant's native timescale, without detection or reconfiguration.

%When a peer's cable severs, the load redistributes within one payload-pendulum period ($\approx 2$~s) and the surviving drone's tension rises. Identity \eqref{eq:Tff} converts that rise at the next control tick into a proportional thrust increment, absorbing the new load at the plant's native rate without detection or communication. The mechanism sits inside a conventional cascade---an outer-loop PD controller with anti-swing slot shift, a single-step convex QP on the tilt-and-thrust envelope, and a geometric attitude loop---with \eqref{eq:Tff} as the sole nonstandard component.

%We develop the consequences of \eqref{eq:Tff} along three axes. The primary contribution (C2, \cref{sec:stability}) is a hybrid practical-ISS theorem: under a single-active-regime QP doctrine and a spaced-fault dwell hypothesis $\tau_d \ge \tau_{\mathrm{pend}}$, the closed-loop tracking error admits a closed-form recovery envelope with Lyapunov decay rate $\lambda$ and per-fault-cycle contraction $\rho \in (0,1)$, conditional on bounded slack excursions, spaced fault events, and a single-active-regime QP---all audited in the simulation campaign. The strict $\rho < 1$ is achieved by \eqref{eq:Tff}: the dwell hypothesis is the minimum time over which $T_i^{\mathrm{ff}} = T_i$ restores the Lyapunov function below its pre-fault level (\cref{lem:dwell-decay}). The first secondary contribution (C1, \cref{sec:reduction-and-reference}) is a taut-cable reduction theorem: on a slack-excursion-bounded domain parameterized by the timescale ratio $\delta$ and duty-cycle bound $\eta$, an explicit first-order error budget justifies replacing the distributed rope dynamics with a lumped effective stiffness; all admissibility gates are met with margin by all six missions (\cref{tab:domain-audit}). The third contribution (C3, \cref{sec:extensions}) is a scalar admissible-window design rule for the $L_1$-adaptive altitude augmentation: two necessary conditions --- discretization stability and bandwidth coverage --- derived from the same scalar discrete pole together place the canonical gain $\Gamma$ inside a certifiable interval $[\Gamma_{\min}, \Gamma^\star]$, reducing reliance on empirical tuning. This is an auxiliary engineering rule, not a full discrete-time $L_1$ stability certificate for the complete interconnection.

%The proposed controller is validated in high-fidelity Drake multibody simulation through a pre-declared six-variant capability demonstration and four targeted stress campaigns on a five-drone, \SI{10}{\kilo\gram}-payload, \SI{4}{\metre\per\second} Dryden-wind scenario under single and compound cable-severance faults. The feed-forward ablation campaign isolates identity \eqref{eq:Tff} as the causal mechanism: disabling it materially increases cruise-tracking error and post-fault altitude sag (\cref{tab:ff-ablation}). Two null findings are reported alongside: in these regimes, a composable extension layer is structurally inactive because the baseline already satisfies the relevant constraint. The architecture admits a sequence of $F \le N - 2$ unannounced severances under the actuator-margin envelope (\cref{sec:fault-and-controller}); the demonstrated $N=5$, $m_L=\SI{10}{\kilo\gram}$ admits $F \le 2$.

%This paper identifies a stability mechanism for structural transitions that self-announces through actuator-side measurements. Four architectural consequences follow: \emph{locality} (no peer state, fault flag, or shared scheduler); \emph{passivity} (no detector, classifier, or reassignment rule); \emph{robustness to $F \le N - 2$ severances} by induction under the actuator-margin envelope; and \emph{team-size independence} (the per-drone law carries no $N$ structurally; dependence concentrates in the actuator-margin envelope). These properties are audited in \cref{sec:discussion-advance}; the two null findings (\cref{rem:mpc-null,rem:reshape-null}) confirm that the baseline mechanism explains the observed behavior.

Cooperative aerial transport uses multiple uncrewed aerial vehicles (UAVs) to carry a shared payload through flexible cables. In practical deployments, cable severance can occur due to abrasion, hardware fatigue, or obstacle contact. A severance event changes the constraint structure of the system, redistributes load across the remaining cables, and requires a compensatory response within one payload-pendulum period. This timescale is on the order of seconds and is shorter than the latency of conventional fault-tolerant control pipelines based on detection, classification, and reconfiguration.

This paper shows that, for this class of structural faults, a simple decentralized feed-forward identity enables passive fault tolerance. Each drone $i$ sets its altitude thrust feed-forward equal to its measured cable tension. 
\begin{equation}
  T_i^{\mathrm{ff}}(t) = T_i(t).
\end{equation}
When a cable severs, the resulting load redistribution appears immediately in the local tension measurement. The controller converts this change into a proportional thrust adjustment at the next control update using only local information. The mechanism requires no fault detection, no communication between agents, and no reconfiguration of the control law.

We analyze the resulting closed loop as a hybrid system with discrete structural transitions and continuous-time dynamics. The feed-forward identity supplies the bounded jump mechanism required for hybrid practical input-to-state stability, which enables recovery across sequences of unannounced cable severances under actuator and dwell-time conditions. The analysis yields an explicit recovery envelope with exponential decay and per-fault-cycle contraction.

We validate the proposed controller in high-fidelity multibody simulation across a range of fault scenarios. The results show that the system absorbs single and multiple cable severances within one payload-pendulum period. Feed-forward ablations support the interpretation that the identity is the active recovery mechanism within the tested baseline architecture, while additional control layers remain inactive on the demonstrated trajectory family.

This subject has not received the attention it deserves. Related work falls into four categories: centralized and distributed cooperative transport, geometric tracking and slung-load control, $L_1$ adaptive control, and fault-tolerant control.

\textit{Centralized and distributed cooperative transport} is a force-allocation architecture~\cite{Michael2011,Fink2011,Tagliabue2019} and distributed predictive variants~\cite{WangOng2017,LiLoianno2023,PallarLi2025,SuBhowmick2023} share a synchronous global-state requirement: the per-tick communication dependency is precisely what a cable-severance transition breaks at the actuator-tick rate, and none of these architectures accounts for cable severance in its stability argument. \textit{Geometric tracking}~\cite{LeeLeokMcClamroch2010}, \textit{flatness-based slung-load}~\cite{Sreenath2013}, and $\mathrm{SE}(3)$ \textit{payload control}~\cite{Lee2018} assume a fixed constraint set; multi-vehicle extensions~\cite{LeeSreenathKumar2013,SharmaSundaram2023} reintroduce global-state exchange, collapsing the strict locality the present architecture requires. \textit{$L_1$ adaptive control} provides matched-disturbance rejection with transient bounds~\cite{CaoHovakimyan2008,Hovakimyan2010book}; its extension to cooperative cable-fault settings under strict locality remains open~\cite{WuCheng2025}.

The fault-tolerant control literature is divided into two main categories: active architectures, which reconfigure themselves after diagnosing a fault~\cite{Blanke2015,ZhangJiang2008}, and passive architectures~\cite{Patton1997,Blanke2015}, which maintain robustness to absorb faults without reconfiguration\cite{LiuSuspensionFailure2021}. In the context of cooperative transport, approaches such as observer-based disturbance estimation with dynamic load reassignment address cable failures by detecting them and reconfiguring the system. Other methods, such as passive force control and leader-follower force estimation~\cite{NoCommTransport2021}, avoid the need to exchange information between UAVs but rely on a fixed number of cables.  Recent work has explored structured representations of system behavior and constraint prioritization as an alternative to reactive or detection-based approaches.~\cite{rulebooks}

%This paper focuses specifically on the processing of structural transition announcements. In our approach, the measurement of tension on the actuator side acts as this announcement. This simplifies the system by eliminating the need for a lengthy process involving detection, classification, and reconfiguration. The key innovation presented here is the stability certification of a decentralized mechanism that converts tension into thrust in the presence of a structural fault. We directly measure the fault's impact via local cable tension.

Existing approaches either rely on detection and reconfiguration or assume a fixed constraint set. None provide a stability certificate for decentralized control under unannounced structural transitions with strict locality. In contrast, this paper identifies a specific feed-forward identity that enables passive fault tolerance and provides a rigorous hybrid stability analysis to certify its effectiveness under explicit conditions. The contribution is not just the mechanism itself, but the demonstration that it can be certified to work across structural transitions without detection or communication.The literatures overlooks the following important aspects: .

\textbf{\textit{Why measured-force feedback and passive force control are insufficient?}} Measured-force feedback and passive impedance/admittance control are mature techniques for modulating stiffness or damping in fixed-topology systems. However, cable severance is not a disturbance on a fixed constraint structure; it is a \emph{structural transition} that changes the cardinality and topology of the constraint set. Conventional passive force-control architectures do not address this class of faults because they assume a fixed number of cables and do not account for the discrete dynamics of a structural transition. Measured tension alone cannot induce a stabilizing response across a structural transition without \emph{explicit certification} that the combined hybrid discrete-continuous dynamics remains stable. The present work addresses this gap: it shows that the identity $T_i^{\mathrm{ff}} = T_i$ is not merely a measured-force gain; it is a \emph{mechanism for bounded-jump stabilization} in a hybrid system. The key contribution is the hybrid practical-ISS certificate that relates the identity to post-fault recovery through the payload-pendulum timescale constraint and the dwell hypothesis $\tau_d \ge \tau_{\mathrm{pend}}$. Without this certificate, the identity is heuristic; with it, the response to cable severance becomes predictable and auditable.

\textbf{\textit{Why strict locality and no-communication decentralization matter?}} The fault-tolerant-control literature is accustomed to active reconfiguration: when a component fails, a centralized or coordinated mechanism detects it, classifies the fault, and reconfigures the control law or resource allocation. This pipeline requires communication and consensus, which takes time—time unavailable on the payload-pendulum period. The present approach abandons the detection-classification-reconfiguration pipeline in favor of a \emph{self-announcing mechanism}: cable severance immediately changes the load distribution, and each drone's local tension sensor observes this change within one or two control ticks. By coupling this local observation directly to thrust, without peer-state exchange or a centralized diagnosis, the system responds passively and with zero latency outside the plant's native bandwidth. This eliminates the temporal gap between fault occurrence and response onset. The trade-off is that strict locality and no-communication prevent global optimization (such as ideal load sharing among survivors) and require a conditional certificate rather than a uniform one. The analysis certifies the decentralized no-communication response under explicit admissibility conditions (H1–H3), which the simulation campaign audits for the demonstrated mission family.


The paper offers three contributions to the current understanding of decentralized fault-tolerant control in cooperative aerial transport:

First, we introduce a strictly local tension-to-thrust feed-forward identity, $T_i^{\mathrm{ff}} = T_i$, for passive cable-severance tolerance in cooperative aerial transport. The identity converts the self-announcing redistribution of cable tension after severance into an immediate thrust correction, without fault detection, inter-agent communication, or controller reconfiguration. A reduced-order taut-cable model with explicit error bounds (\cref{thm:reduction}) supplies the plant on which the identity is analyzed. (C1, \cref{sec:reduction-and-reference})

Second, we prove a conditional hybrid practical input-to-state stability certificate for the resulting decentralized closed loop. The proof combines a slack-excursion-bounded taut-cable reduction, inter-fault ISS decay, bounded fault jumps, and dwell-cycle contraction under explicit actuator and dwell-time conditions, yielding an explicit recovery envelope with exponential decay and per-fault-cycle contraction $\rho \in (0,1)$. (C2, \cref{sec:stability})

Third, we validate the mechanism in high-fidelity multibody simulation under single- and dual-cable severance. Feed-forward ablations show that disabling $T_i^{\mathrm{ff}} = T_i$ inflates cruise RMSE by $34$--$39\%$ and peak post-fault sag by $3.6$--$4.0\times$, supporting mechanism attribution to the identity within the tested baseline architecture. Auxiliary layers ($L_1$, MPC, reshape) preserve the certificate when active and are characterized as non-binding on the demonstrated trajectory family. (C3, \cref{sec:simulation})