The campaign of \cref{sec:simulation} establishes a single architectural commitment: when the structural transition self-announces through actuator-side measurements on the plant's native timescale, the announcement-detection-reconfiguration timeline of classical active fault-tolerant pipelines is unnecessary, and a single decentralized identity~$T_i^{\mathrm{ff}} = T_i$ on the per-drone altitude channel suffices. The five-drone, ten-kilogram dual-fault scenario is one point on the feasibility surface (\cref{fig:feasibility-envelope}). We discuss what the campaign establishes and where its boundary leaves open questions.

\subsection{Audit of the Three Contributions}
\label{sec:discussion-audit}

\Cref{thm:reduction} (C1) is defended in two layers. The domain-gate audit (\cref{tab:domain-audit}) certifies all six missions sit inside $\Omega^{\mathrm{dwell}}_{\tau}$ with non-trivial margin. The fidelity layer (\cref{sec:VI-reduction}) quantifies the per-rope dispersion the lumped scalar abstracts: cruise-window RMS residual $\SI{12.4}{\newton}$ on dual-fault variants, peaks $\SI{60}{\newton}$ at the most centripetally-loaded instants. The dispersion is informative rather than alarming: the $\mathcal{O}(\delta+\eta_{\max})$ guarantee bounds closed-loop trajectory deviation, not per-rope difference, and the identity reads each drone's \emph{own} measured tension. A trajectory-level comparison against a parallel reduced-rope simulator is the principal extension of the C1 audit.

\Cref{thm:hybrid-stability} (C2) is supported by three converging measurements: all six variants pass the pre-registered thresholds with dual-fault peak sag $\SI{95.2}{\milli\metre}$ vs $\SI{100}{\milli\metre}$ ceiling (\cref{tab:performance-ci}); fits on V4/V5 Lyapunov proxy give $\hat\rho_{\mathrm{V4}}=0.20$, $\hat\rho_{\mathrm{V5}}=0.045$; the dwell sweep shows $\hat\rho < 1$ across $\tau_d/\tau_{\mathrm{pend}}\in[0.5,3.0]$ including the sub-threshold point. \Cref{prop:gamma-star} (C3) is validated by the stability dashboard: across $\Gamma$ spanning $2.4$ decades, RMSE is unimodal with smooth degradation past $\Gamma^\star$---the signature of a sufficient-condition Lyapunov bound. The three contributions converge on a single causal anchor: the P2-A ablation (\cref{tab:ff-ablation,fig:self-announcement}) raises cruise RMSE by $34$--$39\%$ and inflates peak sag $3.6$--$4.0\times$, isolating the identity as the mechanism on which C1, C2, and C3 jointly depend.

\subsection{What the Architecture Changes}
\label{sec:discussion-advance}

Existing decentralized cooperative-transport architectures address fault tolerance through one of three families. Active fault-tolerant pipelines~\cite{LiuSuspensionFailure2021} estimate, classify, and reassign on a detection-confirmation-reassignment timeline comparable to or longer than $\tau_{\mathrm{pend}}\approx \SI{2.24}{\second}$. Passive force-control and leader--follower architectures~\cite{GassnerPassiveForce2017,NoCommTransport2021} avoid inter-UAV exchange but presuppose fixed cable cardinality, with no guarantee across the constraint-set transition. Centralized and synchronous-distributed force allocation~\cite{Michael2011,WangOng2017,LiLoianno2023,PallarLi2025,SuBhowmick2023} is structurally inapplicable: the per-tick global-state synchronization it requires is precisely what the constraint-cardinality transition breaks at the actuator-tick rate.

The commitment here is to read the transition off the actuator-side measurement and convert it directly into a thrust increment through $T_i^{\mathrm{ff}}=T_i$, eliminating the announcement-detection-reconfiguration timeline rather than substituting a faster version. The empirical signature is the per-fault recovery times in \cref{tab:recovery-iae}: $\SI{0.00}{\second}$ on first faults (sub-tick) and $\SI{0.32}{\second}$, $\SI{0.33}{\second}$ on V4/V6 second faults, well inside one pendulum period. The four architectural properties of the introduction follow as consequences: \emph{locality} by $\mathcal{I}_i$ \eqref{eq:info-set}; \emph{passivity} by the absence of detector, classifier, or scheduler in the loop; \emph{robustness to multi-agent failure} by induction on the fault counter, with the demonstrated $(N{=}5,m_L{=}\SI{10}{\kilo\gram},F{=}2)$ consuming $27\%$ of the envelope (\cref{fig:feasibility-envelope}); and \emph{team-size independence}, since the per-drone law and lemma stack carry no $N$ structurally and the dependence concentrates in $m_Lg/N$ and $m_Lg/(N{-}F)$.

The two null findings (NF2, NF3) are architectural evidence rather than failed extension claims. The MPC layer shows essentially identical peak tension $\approx \SI{104.6}{\newton}$ across all ceilings (\cref{tab:mpc-ceiling-sweep}) because the pre-fault dynamic spike dominates the quasi-static load the ceiling targets. The reshape supervisor returns $0.0\%$ benefit at every tested period (\cref{rem:reshape-null}) because the dynamic centripetal load dominates the hover asymmetry \cref{thm:reshape} optimises against. Both nulls show the conventional active-FT machinery to be \emph{structurally inactive} on the demonstrated family; the bookkeeping of \cref{sec:theorem-preservation} preserves \cref{thm:hybrid-stability} unchanged.

\subsection{Boundary of the Architectural Category}
\label{sec:discussion-boundary}

Three regimes lie outside the analysis domain. Rapid multi-fault ($\tau_d < \tau_{\mathrm{pend}}$) voids \cref{lem:dwell-decay}; constructing a regime that empirically violates contraction is the natural completion of the dwell-boundary characterisation. Long slack excursions exit $\Omega^{\mathrm{dwell}}_{\tau}$, and QP fallback breaks D1; both are open to the same boundary search and benefit from a multi-fault extension of the dwell argument beyond pairwise composition.

A second class targets the linearisation residuals. Tube-MPC tightening folding the recovery-envelope overshoot into the QP RHS would convert the P2-C null into a positive result on spike-dominated regimes; an aggressive-period reshape with an objective extended to the dynamic centripetal load would similarly close the P2-D null; main-mode $L_1$ evaluation at the canonical $\SI{10}{\kilo\gram}$ point would complete C3-$L_1$; and a four-axis $(m_L, k_s, c_{\mathrm{drag}}, \bar{w})$ response surface on CRN-paired seeds is the visualization counterpart of the same claim. A third class lies outside the architectural category itself---drone-side faults, soft cable failures, tension-sensor failures, payload-attitude dynamics---enumerated in \cref{sec:theorem-preservation}. These are not gaps in the present campaign.

Within this category, the campaign establishes that the identity $T_i^{\mathrm{ff}}=T_i$ delivers hybrid practical input-to-state stability with measured contraction margin across single and dual unannounced severances, that the identity is necessary in the precise sense P2-A isolates, and that the conventional active-FT machinery is structurally inactive on the trajectory family for which the identity is binding.
