
We study cooperative aerial transport under unannounced cable-severance faults and formalize the plant, fault model, admissible controller class, reference trajectory, taut-cable reduction, and formal problem. The system consists of:

\paragraph{Vehicles, payload, and state.} A team of $N$ quadrotor UAVs, indexed $i\in\mathcal{N}\triangleq\{1,\dots,N\}$, operates in an inertial world frame $\mathcal{W}$. Drone $i$ has mass $m_i$, diagonal inertia $\mat J_i$, position $\vect p_i$, velocity $\vect v_i$, attitude $\mat R_i\in\SOthree$, and angular velocity $\vect\omega_i$, with per-drone state $x_i\triangleq(\vect p_i,\vect v_i,\mat R_i,\vect\omega_i)$. A point-mass payload of mass $m_L$ has state $x_L\triangleq(\vect p_L,\vect v_L)$; payload rotation is neglected.

\paragraph{Cables and complementarity condition.} Each drone $i$ is connected to the payload by a flexible cable. Cable $i$ exerts a world-frame force $\vect T_i^{\mathcal{W}} \in \R^3$ on the payload-side attachment point and an equal and opposite reaction on the drone-side attachment point. The cable transmits tensile force only when taut; its scalar tension is
\begin{equation}
 T_i \;\triangleq\; \bigl\|\vect T_i^{\mathcal{W}}\bigr\|
     \;\ge\; 0,
  \label{eq:tension-def}
\end{equation}
with the unilateral constraint $T_i = 0$ whenever the cable is slack. The complementarity condition
\begin{equation}
 T_i \;\ge\; 0, \quad
  \ell_i \;-\; L \;\le\; 0, \quad
 T_i\,\bigl(\ell_i - L\bigr) \;=\; 0,
  \label{eq:complementarity}
\end{equation}
where $\ell_i = \|\vect p_{a, i} - \vect p_{b, i}\|$ is the cable's instantaneous length, $\vect p_{a, i}$ and $\vect p_{b, i}$ are the drone- and payload-side attachment points, and $L$ is the rest length, captures the unilateral tensile constraint for an idealized inextensible cable. In the simulator each cable is a Kelvin--Voigt spring ($L = \SI{1.25}{\metre}$, $k_s = \SI{25000}{\newton\per\metre}$); the elastic correction to the inextensible model is $O(\delta)$ and is absorbed in \cref{thm:reduction}.

\paragraph{Equations of motion.} The drone translational dynamics
\begin{equation}
 m_i\,\dot{\vect v}_i
    = f_i\,\mat R_i\,\hat{\vect e}_3
    - m_i g\,\hat{\vect e}_3
    - \vect T_i^{\mathcal{W}}
    + \vect F_{\mathrm{w},i}(t)
  \label{eq:drone-eom}
\end{equation}
combine thrust $f_i\in[f_{\min},f_{\max}]$, gravity, cable reaction, and wind. Standard rigid-body rotational dynamics $\mat J_i\dot{\vect\omega}_i+\vect\omega_i\times\mat J_i\vect\omega_i=\vect\tau_i$, $\dot{\mat R}_i=\mat R_i[\vect\omega_i]_\times$ govern attitude. The payload obeys
\begin{equation}
 m_L\,\dot{\vect v}_L
    = -m_L g\,\hat{\vect e}_3
    + \sum_{i\in\mathcal{S}(t)} \vect T_i^{\mathcal{W}}
    + \vect F_{\mathrm{w},L}(t),
  \label{eq:payload-eom}
\end{equation}
with $\mathcal{S}(t)\subseteq\mathcal{N}$ the surviving-cable set. Per-drone inputs are $u_i=(f_i,\vect\tau_i)\in\mathcal{U}_i\triangleq[f_{\min},f_{\max}]\times[-\tau_{\max},\tau_{\max}]^3$, the plant state $x=(x_{\mathcal{N}},x_L)$ lives in $\mathcal{X}$, and the admissible disturbance set is
\begin{equation}
\label{eq:disturbance-set}
  \mathcal{D} \triangleq \bigl\{ d(\cdot) :
    \|\vect F_{\mathrm{w},i}(t)\|,\; \|\vect F_{\mathrm{w},L}(t)\|
      \le W_{\max},\;\forall i,t \bigr\},
\end{equation}
with $W_{\max}=\SI{1}{\newton}$; the Dryden realization used in V2--V6 stays inside this bound (\cref{tab:sim-params}).

\input{Figures/diagrams/fig_physical_setup.tex}

At a time $t_k^\star$, cable $i_k^\star \in \mathcal{N}$ severs instantaneously. This structural fault event permanently removes cable $i_k^\star$ from the constraint set:
\begin{equation}
 T_{i_k^\star}(t) \;=\; 0
  \qquad \forall\, t \ge t_k^\star.
  \label{eq:fault-tension-zero}
\end{equation}
The active cable set transitions discretely:
\begin{equation}
  \mathcal{S}(t_k^{\star,+}) \;=\; \mathcal{S}(t_k^{\star,-})
    \;\setminus\; \{i_k^\star\},
  \label{eq:S-transition}
\end{equation}
reducing the cardinality from $|\mathcal{S}|$ to $|\mathcal{S}| - 1$. For an elastic cable, the kinematic state $(\vect p_i, \vect v_i, \mat R_i, \vect\omega_i, \vect p_L, \vect v_L)$ is continuous across severance; only the ODE right-hand side changes discontinuously at $t_k^\star$.

We consider a sequence of $F \le N-2$ fault events at times $\{t_k^\star\}_{k=1}^F$ with $t_1^\star < \cdots < t_F^\star$. The bound $F \le N-2$ keeps at least two cables intact; actuator margins limit the demonstrated scenario ($N = 5$, $m_L = \SI{10}{\kilo\gram}$) to $F \le 2$. The inter-fault interval $\tau_{d,k} \triangleq t_{k+1}^\star - t_k^\star$ is the \emph{dwell time}. We assume each fault is separated from the previous by at least one payload-pendulum period:
\begin{equation}
  \tau_{d,k} \;\ge\; \tau_{\mathrm{pend}} \;\triangleq\;
    2\pi\sqrt{L/g}, \qquad k = 1, \dots, F-1,
  \label{eq:dwell-assumption}
\end{equation}
reflecting the physical requirement that load redistribution from fault $k$ settle before fault $k+1$. At $L = \SI{1.25}{\metre}$ and $g = \SI{9.81}{\metre\per\second\squared}$, $\tau_{\mathrm{pend}} \approx \SI{2.24}{\second}$. The admissible fault-schedule set is
\begin{align}
\label{eq:fault-schedule-set}
  \mathcal{F} \;\triangleq\;
    \Bigl\{
      \{(t_k^\star, i_k^\star)\}_{k=1}^F :
 &F \le N-2,\;\\
 &t_{k+1}^\star - t_k^\star \ge \tau_{\mathrm{pend}},\;
 i_k^\star \in \mathcal{S}(t_k^{\star,-})
    \Bigr\},\nonumber
\end{align}
$F = 0$ corresponds to the no-fault case.

\paragraph{Admissible controller class.} No fault flag, annunciation, or detection signal is transmitted; the fault manifests through a rise in $T_i(t)$ and a transient payload-velocity deviation, both in $\mathcal{I}_i(t)$. Field-deployable avionics impose a strict \emph{admissible information pattern}. Drone $i$'s controller $\pi_i$ is restricted to the information set
\begin{equation}
  \mathcal{I}_i(t) \;\triangleq\;
    \bigl\{x_i(t),\; T_i(t),\; \vect v_L(t),\;
      \vect p_L^d(\cdot),\; \vect v_L^d(\cdot)\bigr\},
  \label{eq:info-set}
\end{equation}
comprising the drone's inertial state $x_i$, its rope-tension measurement $T_i$, the payload velocity $\vect v_L$ (measured via a co-located downward-facing optical-flow sensor, no drone-to-drone communication), and a pre-distributed reference $({\vect p}_L^d, {\vect v}_L^d)$ loaded before flight. Payload position $\vect p_L$ is deliberately excluded; only velocity enters the closed loop, via the anti-swing slot shift of \cref{sec:proposed}. Each $\pi_i$ is a causal feedback map,
\begin{equation}
 u_i(t) \;=\; \pi_i\!\left(
    \left.\mathcal{I}_i\right|_{[0,t]}
  \right) \in \mathcal{U}_i,
  \label{eq:controller-map}
\end{equation}
possibly with a finite-dimensional internal state, but never with access to information outside \eqref{eq:info-set}. The admissible controller class is denoted $\Pi_{\mathrm{loc}}$. $\mathcal{I}_i(t)$ excludes peer state, peer tension, fault flags, and run-time inter-drone packets; each $\pi_i$ uses only co-located sensors and a pre-flight plan.

\subsection{Reference Trajectory and Taut-Cable Reduction}
\label{sec:reduction-and-reference}

The desired payload motion is a three-dimensional lemniscate trajectory with a sinusoidal altitude variation,
\begin{align}
  \label{eq:lemniscate}
  \vect p_L^d(t) \;=\;
    \begin{bmatrix}
      \dfrac{a\cos\phi(t)}{1 + \sin^2\!\phi(t)} \\[6pt]
      \dfrac{a\sin\phi(t)\cos\phi(t)}{1 + \sin^2\!\phi(t)} \\[6pt]
 z_0 + h_z\sin(\phi(t))
    \end{bmatrix},
  \phi(t) = \frac{2\pi(t - t_0)}{T_{\mathrm{ref}}},
\end{align}
where $a = \SI{3.0}{\metre}$ is the horizontal scale, $T_{\mathrm{ref}} = \SI{12}{\second}$ the period, $z_0 = \SI{3.0}{\metre}$ the cruise altitude, and $h_z = \SI{0.35}{\metre}$ the altitude excursion amplitude. These parameters yield a peak centripetal load of $\approx 0.25g$ at the figure-8 tips.

Each rope of rest length $L$ is discretized into $N_b = 8$ identical Kelvin--Voigt beads of mass $m_{\mathrm{bead}}$ connected by $N_{\mathrm{seg}} = N_b + 1 = 9$ segments of stiffness $k_s$ and damping $c_s$; $N_{\mathrm{seg}}$ is reserved for rope segments and is distinct from the surviving-cable cardinality $N_s(t) = |\mathcal{S}(t)| = N - F(t)$ used in \cref{sec:stability}. The distributed dynamics evolve on a fast timescale,
\begin{equation}
  \tau_{\mathrm{rope}}
    \;=\; 2\pi\sqrt{m_{\mathrm{bead}}/k_s}
    \;\approx\; \SI{6.3}{\milli\second},
  \label{eq:tau-rope}
\end{equation}
separated from the payload-pendulum timescale $\tau_{\mathrm{pend}} \approx \SI{2.24}{\second}$ by $\delta = \tau_{\mathrm{rope}} / \tau_{\mathrm{pend}} \approx 0.003$. On the slow manifold, the net tensile force at the drone-side attachment equals the scalar $\xi_i = k_{\mathrm{eff}}(\ell_i - L)^+$, with effective stiffness
\begin{equation}
 k_{\mathrm{eff}}
    \;=\; \frac{k_s}{N_{\mathrm{seg}}}
    \;=\; \frac{\SI{25000}{\newton\per\metre}}{9}
    \;\approx\; \SI{2778}{\newton\per\metre}.
  \label{eq:keff}
\end{equation}
The six pre-registered missions (V1--V6) keep the system inside $\Omega_\tau^{\mathrm{dwell}}$ with margin (worst slack run $\SI{35.8}{\milli\second}$, worst duty cycle $1.66\%$, \cref{tab:domain-audit}). The reduction replaces the $3N_b$-dimensional bead-chain state with the scalar $\xi_i$; the per-instant taut error is $O(\delta)$ and slack excursions add $O(\eta_{\max})$.

\begin{theorem}[Taut-cable reduction]
\label{thm:reduction}
On the slack-excursion-bounded domain
\begin{align}
\label{eq:omega-dwell}
  \Omega_\tau^{\mathrm{dwell}} \;\triangleq\;
    \bigl\{ x \in \mathcal{X} :
      &\text{every slack run} \le \tau_{\mathrm{slack,max}},\;\\
      &\text{duty cycle} \le \eta_{\max} \bigr\},\nonumber
\end{align}
with $\tau_{\mathrm{slack,max}} = \SI{40}{\milli\second}$ and $\eta_{\max} = 2.5\%$, the distributed Kelvin--Voigt rope dynamics are approximated by the lumped scalar model with closed-loop shape-mode deviation
\begin{equation}
  \|T_i^{\mathrm{true}} - T_i^{\mathrm{qs}}\|
    \;\le\; C_1\,\delta + C_2\,\eta_{\max},
  \label{eq:reduction-error}
\end{equation}
for constants $C_1, C_2 > 0$ depending on the bead-chain parameters $(m_{\mathrm{bead}}, c_s, k_s, N_b)$ and the wind bound $W_{\max}$, where $T_i^{\mathrm{qs}} \triangleq k_{\mathrm{eff}}(\ell_i(t) - L)^+$ is the\emph{ rope-specific} quasi-static tension evaluated at the \emph{same} current chord length $\ell_i(t)$ as $T_i^{\mathrm{true}}$.  This bound is on the shape-mode deviation of rope $i$ from its own quasi-static approximation; it is distinct from the inter-rope tension asymmetry $\varepsilon_i = T_i - T_{\mathrm{eff}}$ reported in the simulation audit (\cref{sec:VI-reduction}), which is driven by formation geometry and can be $O(10)$\,N. The reduced-order model is therefore admissible as the plant for the stability analysis of \cref{sec:stability}.
\end{theorem}

\begin{proof}[Sketch]
The proof has two parts corresponding to the two terms in \eqref{eq:reduction-error}.

\noindent\textbf{Part 1 ($C_1\delta$, taut intervals).} Let $\tilde{\mathbf{q}}\in\R^{N_b}$ collect bead deviations from the quasi-static slow-manifold equilibrium $\bar{\mathbf{q}}(t)$ along the rope chord. Linearization yields the shape-mode dynamics $M_b\ddot{\tilde{\mathbf{q}}}+C_b\dot{\tilde{\mathbf{q}}}+K_b\tilde{\mathbf{q}}=-M_b\ddot{\bar{\mathbf{q}}}$ with $C_b,K_b$ the chain damping and stiffness matrices. Differentiating the shape-mode energy $W=\tfrac12\dot{\tilde{\mathbf{q}}}^\top M_b\dot{\tilde{\mathbf{q}}}+\tfrac12\tilde{\mathbf{q}}^\top K_b\tilde{\mathbf{q}}$ and applying Young's inequality gives $\dot W\le -\alpha_b W + D_b$ with $\alpha_b=\lambda_{\min}(C_b)/m_{\mathrm{bead}}$ and $D_b=O(\delta^2)$ via the Tikhonov scaling $\|\ddot{\bar{\mathbf{q}}}\|=O(\delta)$. Gronwall yields $W^\infty=D_b/\alpha_b=O(\delta^2)$; the drone-side tension deviation $T_i^{\mathrm{true}}-T_i^{\mathrm{qs}}=k_s\mathbf{e}_1^\top\tilde{\mathbf{q}}$ then satisfies $\|T_i^{\mathrm{true}}-T_i^{\mathrm{qs}}\|\le C_1\delta\approx\SI{0.072}{\newton}$ after the $\sim\SI{6}{\milli\second}$ transient. The slowest mode has $\zeta=1.2$, $s_1\approx-\SI{538}{\second^{-1}}$.

\noindent\textbf{Part 2 ($C_2\eta_{\max}$, slack intervals).} On any slack run of duration $\le\tau_{\mathrm{slack,max}}$, the chord shortens by $\Delta\ell_{\mathrm{sag}}$ and the drone-side end segment must re-elongate on re-tension, producing a re-tension overshoot bounded by $k_{\mathrm{eff}}\Delta\ell_{\mathrm{sag}}$. Aggregating over the slack-duty budget gives $C_2\approx\SI{0.22}{\newton}$. At the worst observed $\eta_{\max}=1.66\%$, the combined bound is $C_1\delta+C_2\eta_{\max}\approx\SI{0.076}{\newton}$ per rope ($<0.4\%$ of nominal per-drone tension). Both constants depend only on $(m_{\mathrm{bead}},c_s,k_s,N_b)$ and $W_{\max}$. Detailed Gronwall arithmetic and shape-mode constants are in the supplementary material.
\end{proof}

\Cref{tab:domain-audit} confirms all six missions reside inside $\Omega_\tau^{\mathrm{dwell}}$ with margin on every gate: worst slack run $\SI{35.8}{\milli\second}$ ($\SI{4.2}{\milli\second}$ headroom), duty cycle $1.66\%$ ($0.85$~pp headroom), QP-transition fraction $0.10\%$. The theorem is asserted on the demonstrated regime, not an untested all-taut idealization.

The per-drone controller and the lemma stack in \cref{sec:stability} are structurally independent of $N$. The team size enters only through the post-fault redistributed load
\begin{equation}
  T^{\mathrm{post}}(N,F) = \frac{m_L g}{N-F},
   \label{eq:tension-postfault-N}
\end{equation}
yielding the actuator-margin condition
\begin{equation}
  \frac{m_L g}{N-F} \le \kappa_{\mathrm{act}}\,f_{\max},
   \label{eq:actuator-envelope}
\end{equation}
with $\kappa_{\mathrm{act}}\in(0,1)$ reserving headroom for tilt projection, anti-swing residual, and dynamic centripetal load. At canonical $\kappa_{\mathrm{act}}=0.82$, $f_{\max}=\SI{150}{\newton}$, and the demonstrated $(N,m_L,F)=(5,\SI{10}{\kilo\gram},2)$, \eqref{eq:actuator-envelope} reads $\SI{32.7}{\newton}\le\SI{123}{\newton}$, i.e.\ $27\%$ of envelope. The certificate of \cref{sec:stability} carries unchanged at any $(N,m_L,F)$ satisfying \eqref{eq:actuator-envelope}.
\label{sec:formal-problem}
% ----------------------------------------------------------------

With the plant, fault model, reference trajectory, and reduction in place, we state the formal problem. Unilateral cable coupling \eqref{eq:complementarity}, discrete structural fault transitions \eqref{eq:fault-tension-zero}--\eqref{eq:S-transition}, admissible disturbances \eqref{eq:disturbance-set}, and strict locality \eqref{eq:info-set} together yield a hybrid decentralized control problem outside standard cooperative-transport formulations. The design variable is not a centralized control sequence but a family of local causal feedback laws. The problem, posed on the reduced-order model certified by \cref{thm:reduction}, is as follows.

\begin{definition}[Fault-tolerant decentralized aerial transport]
\label{def:problem}
Given the plant \eqref{eq:drone-eom}--\eqref{eq:payload-eom}, the complementarity constraint \eqref{eq:complementarity}, the mission reference \eqref{eq:lemniscate}, all $d(\cdot)\in\mathcal{D}$ \eqref{eq:disturbance-set}, and all $\sigma\in\mathcal{F}$ \eqref{eq:fault-schedule-set}, find a causal controller family $\{\pi_i\}_{i=1}^N\in\Pi_{\mathrm{loc}}^N$ such that the closed loop satisfies the following.
\begin{enumerate}[label=\textbf{P\arabic*},leftmargin=2.0em]
  \item \textbf{Nominal tracking.} The payload tracking error $e_L(t)\triangleq\vect p_L(t)-\vect p_L^d(t)$ is ultimately practically bounded,
    $\limsup_{t\to\infty}\|e_L(t)\|\le\bar e_\infty+\bar e_0(W_{\max})$, with $\bar e_0\in\mathcal{K}$.
    \label{eq:nominal-practical-bound}
  \item \textbf{Passive fault absorption.} After any fault $t_k^\star$ obeying \eqref{eq:dwell-assumption}, $e_L$ recovers to a bounded envelope,
    \begin{align}
    \label{eq:recovery-bound}
      \|e_L(t)\| \le
        &\beta\bigl(\|e_L(t_k^{\star,-})\|,\,t-t_k^\star\bigr)
         + \Gamma_f\,e^{-\lambda(t-t_k^\star)}\\
        &+ \bar e_\infty + \bar e_0(W_{\max}),\quad t\ge t_k^\star,\nonumber
    \end{align}
    for some $\beta\in\mathcal{KL}$ and constants $\Gamma_f,\lambda>0$, with per-fault-cycle contraction $V(x(t_{k+1}^{\star,-}))\le\rho V(x(t_k^{\star,-}))+c$ for $\rho\in(0,1)$ and explicit $c>0$.
  \item \textbf{Actuator feasibility.} $f_i(t)\in[f_{\min},f_{\max}]$ and $\vect\tau_i(t)\in[-\tau_{\max},\tau_{\max}]^3$ for all $t\ge 0$ and all admissible fault sequences, with non-zero margin.
\end{enumerate}
\end{definition}

The discrete drop in $|\mathcal{S}|$ at each severance, the non-smooth taut-to-slack transitions inside $\Omega_\tau^{\mathrm{dwell}}$, and the strict locality constraint together require a hybrid Lyapunov argument on the tension mismatch
\begin{equation}
  \vect\varepsilon_T(t) \;\triangleq\; \vect T(t) - \vect T^{\mathrm{qs}}(t),
  \label{eq:tension-error-state}
\end{equation}
developed in \cref{sec:stability}.


