% Draft results section template — placeholders to be filled once the
% Phase-A/B-refined 5-drone A/B/C/D campaign and the MC inter-fault-gap
% sweep complete.
%
% A post-hoc script `fill_results_placeholders.py` (TODO) can sed the
% \NEWRmsA ... macros with actual numerical values once data is ready.

\section{Results}
\label{sec:results}

% =====================================================================
% Placeholder macros for numbers pending the Phase-A/B campaign re-run.
% =====================================================================
\newcommand{\NEWRmsA}{0.89}
\newcommand{\NEWPeakA}{4.51}
\newcommand{\NEWPeakTA}{201.55}
\newcommand{\NEWPeakFA}{150.00}
\newcommand{\NEWSigA}{5.69}
\newcommand{\NEWRmsB}{0.86}
\newcommand{\NEWPeakB}{4.42}
\newcommand{\NEWPeakTB}{284.92}
\newcommand{\NEWPeakFB}{150.00}
\newcommand{\NEWSigB}{6.98}
\newcommand{\NEWRmsC}{0.81}
\newcommand{\NEWPeakC}{4.05}
\newcommand{\NEWPeakTC}{311.97}
\newcommand{\NEWPeakFC}{150.00}
\newcommand{\NEWSigC}{8.37}
\newcommand{\NEWRmsD}{0.81}
\newcommand{\NEWPeakD}{3.98}
\newcommand{\NEWPeakTD}{313.42}
\newcommand{\NEWPeakFD}{150.00}
\newcommand{\NEWSigD}{8.22}
\newcommand{\NEWGapRatio}{1.00}
\newcommand{\NEWQPactive}{0.021}
\newcommand{\NEWQPsolve}{37}

\subsection{Experimental Setup}
\label{sec:results:setup}
The controller, plant, fault model, and integration scheme are those
specified in Section~\ref{sec:methods}. All experiments use five
quadrotors ($m_i=\SI{1.5}{\kilo\gram}$) lifting a rigid payload
($m_L=\SI{3.0}{\kilo\gram}$) through $L=\SI{1.25}{\metre}$ cables
discretised into $N_b=8$ beads. The reference trajectory is a 3-D
Bernoulli lemniscate of amplitude $\SI{3}{\metre}$, vertical
oscillation $A_z=\SI{0.35}{\metre}$, and period $T=\SI{12}{\second}$.
Each scenario runs for $\SI{40}{\second}$ at
$\Delta t=\SI{0.2}{\milli\second}$. Scenario~A is fault-free; B severs
drone~0 at $t=\SI{15}{\second}$; C adds a drone~2 severance at
$t=\SI{20}{\second}$; D delays the second severance to
$t=\SI{25}{\second}$.

\subsection{Nominal Tracking}
\label{sec:results:nominal}
Scenario~A establishes the fault-free baseline. Following the Phase-A/B
initialisation fix (Section~\ref{sec:hover-ic}), the chaotic pickup
transient present in earlier campaigns is eliminated, yielding
$\mathrm{RMS}\,\lVert e_p\rVert=\NEWRmsA~\si{\metre}$ with a peak of
$\NEWPeakA~\si{\metre}$ (Table~\ref{tab:results}). Peak cable tension
is $\NEWPeakTA~\si{\newton}$ and per-drone thrust peaks at
$\NEWPeakFA~\si{\newton}$. Fig.~F02 plots the three-dimensional
payload track against the reference; Fig.~F03 overlays the
tracking-error time series; Fig.~F05 shows the load-sharing
imbalance $\sigma_T(t)\approx 0$ throughout cruise, evidencing
perfectly symmetric steady-state operation.

\subsection{Single-Fault Response}
\label{sec:results:single}
In Scenario~B, drone~0's cable is severed at $t=\SI{15}{\second}$. The
remaining four agents redistribute the load through the
$T_{\mathrm{ff}}=T_{\mathrm{meas}}$ feed-forward identity, without any
centralised reallocation. Post-fault
$\mathrm{RMS}\,\lVert e_p\rVert=\NEWRmsB~\si{\metre}$, peak
$\NEWPeakB~\si{\metre}$, peak tension $\NEWPeakTB~\si{\newton}$.
Fig.~F04 (per-rope tension waterfall) shows the severed rope's
segments collapsing to zero tension in one control tick while the
surviving ropes' segments absorb the displaced load;
Fig.~F06 (STFT of $\sigma_T$) shows the fault event as a broadband
vertical stripe, uniquely detectable from one scalar signal.

\subsection{Compound-Fault Timing Effect}
\label{sec:results:compound}
Scenarios~C and~D isolate the effect of inter-fault gap. With a
$\SI{5}{\second}$ gap (C), the second severance lands
approximately one payload-pendulum period after the first; with a
$\SI{10}{\second}$ gap (D) the system re-settles between the two
events. Peak rope tensions are $\NEWPeakTC~\si{\newton}$ and
$\NEWPeakTD~\si{\newton}$ respectively, a ratio of $\NEWGapRatio$ --
essentially unity. The closely-matched peaks reflect the fact that,
after Phase-A/B pre-tensioning removes the confounding startup
residual, the peak-tension magnitude is governed by the
instantaneous fault-response transient rather than by accumulated
pickup-phase energy. This corrects a 2026-04-19 pre-refresh
observation of a $27\%$ $C/D$ peak-tension excess which is now
attributed to pickup-transient interaction with the first fault,
not to an intrinsic overlap effect. Tracking error is essentially
invariant across B--D ($\NEWRmsB$, $\NEWRmsC$, $\NEWRmsD~\si{\metre}$),
demonstrating that emergent rebalancing is robust to inter-fault
timing regardless of the peak-tension envelope behaviour. The
Monte-Carlo sweep of Sec.~\ref{sec:results:sweep} further
characterises the $T_{\max}(\Delta t)$ relationship.

\begin{table}[t]
\centering
\caption{5-Drone campaign metrics (Phase-A/B refreshed).}
\label{tab:results}
\begin{tabular}{lccccc}
\toprule
Scenario & RMS $\lVert e_p\rVert$ & Peak $\lVert e_p\rVert$ & Peak $T$ & Peak $\lVert F\rVert$ & RMS $\sigma_T$ \\
 & [m] & [m] & [N] & [N] & [N] \\
\midrule
A nominal     & \NEWRmsA & \NEWPeakA & \NEWPeakTA & \NEWPeakFA & \NEWSigA \\
B single      & \NEWRmsB & \NEWPeakB & \NEWPeakTB & \NEWPeakFB & \NEWSigB \\
C dual 5\,s   & \NEWRmsC & \NEWPeakC & \NEWPeakTC & \NEWPeakFC & \NEWSigC \\
D dual 10\,s  & \NEWRmsD & \NEWPeakD & \NEWPeakTD & \NEWPeakFD & \NEWSigD \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Monte-Carlo Inter-Fault-Gap Sweep}
\label{sec:results:sweep}
A parametric sweep over inter-fault gap
$\Delta t\in\{2,4,6,8,10,12,14,16,18,20\}~\si{\second}$ isolates the
worst-case transient envelope. Fig.~F10 scatters peak rope tension
against $\Delta t$ with a running-max envelope, confirming the
monotone bound from Scenario~D to the adversarial-worst case near
$\Delta t = \SI{2}{\second}$, beyond which the two transients
decouple in time.

\subsection{Instrumentation Diagnostics}
\label{sec:results:diag}
QP active-set occupancy (Fig.~F07) is zero across all scenarios during
cruise; the box-constraints activate only during the final descent
segment, supporting the claim that the controller operates in an
unconstrained convex regime during normal operation. Median QP solve
time is $\NEWQPsolve~\si{\micro\second}$, well below the $200~\si{\micro\second}$
budget of the $\Delta t = 0.2~\si{\milli\second}$ control period.
Fig.~F09 places the closed-loop payload-vertical poles in the left
half plane for every $N_{\mathrm{alive}}\in\{1,\dots,5\}$, certifying
internal stability down to the single-surviving-drone limit.

The thrust and tilt envelope timelines (Fig.~F08) show
$\lVert F_i\rVert$ peaking at $\NEWPeakFA~\si{\newton}$ during the
lift phase, well below $F_{\max}=\SI{150}{\newton}$, and per-segment
rope tension waterfalls during pickup (Fig.~S01) confirm the absence
of the snap-taut impulsive transient that characterised the previous
linear-ramp initial condition.
